---
title: "In Class Assignment 5"
author: "Christopher Hainzl, Christopher Hakkenberg, Christopher Barbieri, Ryan Podzielny, Peter Bitanga"
date: "2022-11-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
```

```{r}
data <- read.table("table12-16.txt", sep = ",", header = TRUE)
head(data)
```

```{r}
str(data)
```

```{r}
data$Hth <- as.factor(data$Hth)
```

```{r}
str(data)
```

```{r}
colSums(is.na(data))
```

```{r}
table(data$Hth)
```

```{r}
prop.table(table(data$Hth))
```

```{r}
logistic.model <- glm(Hth~Fib + gam, family = "binomial", data = data)
summary(logistic.model)
```

# Part A

Fib is related to the probability that a patient has an unhealthy level of ESR, and gam isn't. Fib has a p-value of 0.0491 (which is less than 0.05), and gam has a p-value of 0.1925 (which is greater than 0.05).

# Part B
Beta 0 is -12.7921, Beta 1 is 1.9104 (Fib), and Beta 2 is 0.1558 (gam).

```{r}
confint(logistic.model, level = 0.95)
```

Based on the values above, we are 95% confident that the estimates are within those ranges.

# Part C
```{r}
prob_esr <- function(x, y) {
  eee <- exp((-12.7921+(1.9104 * x) + (0.1558 * y)))
  eee/(1+eee)
}
```

```{r}
prob_esr(2.52, 38)
```

Group 3: 
Christopher Hainzl, Christopher Hakkenberg, Christopher Barbieri, Ryan Podzielny, Peter Bitanga