---
title: "Homework 3"
author: "Christopher Hainzl"
date: "2022-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r packages}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## Question 1


```{r}
baseball <- read.table(file=
"http://jse.amstat.org/datasets/baseball.dat.txt",
header = F, col.names=c("salary", "batting.avg", "OBP", "runs", "hits",
"doubles", "triples", "homeruns", "RBI", "walks", "strike.outs",
"stolen. bases", "errors", "free.agency.elig", "free.agent.91",
"arb.elig", "arb.91", "name"))
head(baseball)
```
## 1a

```{r}
colSums(is.na(baseball))
```
- No, there isn't any missing data.

```{r}
# Determine the variable types of the columns
sapply(baseball, class)
```
- 4 predictors are categorical (free.agency.elig, free.agent.91, arb.elig, arb.91), and 10 predictors are continuous, integer counts.

## 1b

```{r}
# Histogram of salaries without log transformation
ggplot(data = baseball, mapping = aes(x = salary)) +
  geom_histogram(bins = 50, fill = "yellow") +
  labs(title = "Salary Histogram (w/o log transformation)")
```

```{r}
# perform log transformation on salary
baseball[1] <- log(baseball[1])
```

```{r}
# Histogram of salaries with log transformation
ggplot(data = baseball, mapping = aes(x = salary)) +
  geom_histogram(bins = 50, fill = "red") +
  labs(title = "Salary Histogram (with log transformation)")
```

- The histogram without the log transformation appears to be right-skewed, while the histogram with the log transformation does not appear to have any sort of skewness.

## Question 2

## 2a

```{r}
set.seed(123)
# Sample Indexes
Index = sample(1:nrow(baseball), size = 0.7*nrow(baseball))
# Splitting Data
TrainData = baseball[Index,]
dim(TrainData)
TestData = baseball[-Index,]
dim(TestData)
```

```{r}
head(TrainData)
```

```{r}
head(TestData)
```

## 2b and 2c

```{r}
# 2b
fit.train <- lm(salary ~ batting.avg + OBP + 
runs + hits + doubles + triples + homeruns + RBI + walks + strike.outs + stolen..bases + errors, 
data = TrainData)

# 2c
summary(fit.train)
```
## 2d
- When working with a 5% significance level and looking at the individual p-values, salary only appears to have a relationship with hits, home runs, and strikeouts (since the p-values for those three predictors are less than 0.05).

## 2e

```{r}
# Use the model for the training data to predict the 
# test data, and output mean squared error
mean((TestData$salary - predict.lm(fit.train, TestData))^2)
```